Discovery of s-process stars in the LAMOST survey

### 2018-01-25 article/ms.tex
draft complete

\begin{abstract}
Stars with peculiar enhancements of carbon and heavy elements ($Z > 30$) are historically referred to as `Barium stars'. While this chemical abundance pattern can be explained by thermally pulsing asymptotic giant branch (AGB) stars, many so-called Barium stars have not evolved through the red giant phase and therefore cannot be AGB stars. These less-evolved stars are generally thought to have accreted mass from a more-evolved AGB companion. For this reason the frequency and properties of less-evolved Barium stars are uniquely informative on the binary fraction, AGB yields, mass transfer, and stellar evolution. Here we present the discovery of 967 s-process candidates from 454,180 giant stars observed by LAMOST. Our sample includes; 257 Barium stars, 49 stars exhibiting only significant barium enhancement, and 661 exhibiting only strontium enhancement. This sample constitutes the largest number of s-process enhanced stars ever discovered, and represents the most comprehensive search for Barium stars in the Milky Way. All of our Barium stars are consistent with membership in the Milky Way disk, and we find an \todo{steady/increasing/decreasing} fraction of Barium stars with stellar metallicity, as \todo{expected} from AGB yields and stellar evolution theory. We find up to 51 Barium stars exhibit some evidence of technetium from LAMOST spectra, and most (46/51) have $\log{g} > 2$, which suggests they cannot be AGB stars and may indicate a recent ($<1\,\textrm{Myr}$) mass accretion event. Most (178/257) Barium stars in our sample show strong carbon enhancement, consistent with the literature. However, despite previous claims indicating that Barium stars are preferentially enhanced in sodium due to the NeNa nucleosynthesis cycle occurring in massive AGB stars, only 5 ($<2$\,\%) of our Barium stars show any evidence of enhancement in sodium. \todo{A sentence about AGB yield comparison.} When binary properties become available from end-of-mission Gaia astrometry, this sample of s-process stars will be critical in explaining mass transfer in binary systems and refining AGB yields across a range of masses and metallicities.
\end{abstract}

@@

\subsection{Extrinsic and intrinsic Barium stars}
Barium stars exist as either intrinsic or extrinsic objects. Intrinsic Barium stars must be massive enough to reach the TP-AGB phase, and extrinsic Barium stars must be in a binary with a previously polluting TP-AGB star companion. Figure \ref{fig:figure2} shows that the majority of our Barium stars have $\log{g} > 2$, highlighting that most stars in our sample are not massive enough to have already reached the TP-AGB phase and therefore, must be extrinsic Barium stars. The classification for Barium stars with $\log{g} < 2$ is uncertain: these stars could be intrinsic or extrinsic. From here on when we refer to extrinsic Barium stars, we specifically refer to those with $\log{g} > 2$.

\subsection{Sodium enhancment}
According to \citet{el1995}, sodium overabundances observed in the atmospheres of supergiants is synthesised by the NeNa reaction chain in the convective core of main-sequence stars. Sodium is transported to the surface of these supergiants via mixing of CNO cycle products during the first dredge-up, although, it is important to note sodium enhancement is not constrained to supergiant stars. Stars with a  minimum mass of $1.5\,M_\odot$ can also exhibit enhanced sodium abundances through the NeNa cycle \citep{denissenkov1987}. \citet{antipova2004} reported sodium enhancements in 16\todo{/??} of their Barium star candidates, and they relate this to the dredge-up of nuclear-burning material produced by convection during the red-giant phase. Moreover, they suggest [Na/Fe] ratios exhibit systematically higher values for giants with lower $\log{g}$ values, but do not exhibit variations in metallicity. Similarly, \citet{decastro2016} highlights the possible weak anti-correlation between [Na/Fe] ratio and $\log{g}$, and highlights that this trend in previous studies \citep[e.g.][]{boyarchuk2002,mishenina2006,luck2007,takeda2008}.

We find only $<2$\,\% of our Barium stars show sodium enhancement, contrary to previous studies \citep[e.g.][]{decastro2016}. We also note that the 5 sodium-enhanced Barium stars all have $\log{g} \approx 3$, not at lower $\log{g}$ values as might be expected if low $\log{g}$ values systematically contributed to higher [Na/Fe] abundance ratios. These results suggest that sodium-rich material from a companion TP-AGB star has polluted these 5 Barium candidates, but importantly, sodium enhancement is by no means ubiquitous in Barium stars. The comparisons made in \citep{decastro2016} made use of multiple literature comparisons, and therefore it is possible that there are systematic differences (biases) between Milky Way studies of normal FGK-type stars and those focussed on Barium stars, leading to unresolved differences in the reported [Na/Fe] abundance ratio. These same systematic effects could contribute to the observed correlation between [Na/Fe] and $\log{g}$, which we do not observe in our sample.

% Could be removed:
%Also, synonymous to our s-process candidates, synthesised sodium is brought to the surface of giants via the first dredge-up, therefore, these candidates may display sodium enhancement without mass transfer from a TP-AGB star.

%Furthermore, this suggests other enhancements including our 5 sodium and 178 carbon enhanced Barium stars exist through TP-AGB companion pollution.

\subsection{Technetium enhancement}
\todo{Technetium has a half life of approximately 211,000 years and as a result cannot be present in an extrinsic Barium star. Enhancemnt in Tc would indicate the presence of AGB stars \citep{jorissen1993}, and suggest some of our candidates are intrinsic Barium stars. Our analyse discovered 51 single Tc line matches within our 257 Barium star candidate sample. However, for Tc to be present in  a star the enhancement lines must be considerably strong, and given that all 51 matches were at single lines it is concluded that these results are data artifacts. In agreement with previous findings \citep[e.g.][]{little1987,smith1984,smith1983}, the results indicate that none of the evolved Barium stars in our sample exhibited the presence of Tc enhancement, and it can be concluded our Barium star sample is extrinsic. It is noted, Tc enhancement is a very difficult line to analyse and that this conclusion is not concrete.}

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{HRdummy.png}
    \caption{HR Diagram}
    \label{fig:figure2}
\end{figure}

\subsection{AGB yields comparison}

@@

Our research discovered 967 s-process candidates including; 257 Barium stars, 49 stars exhibiting exclusively barium enhancement, and 661 exhibiting exclusively strontium enhancement from 454,180 giant stars observed by LAMOST. This is the largest sample of s-process enhanced stars to date \todo{by a factor of how many? what's the next largest size sample?}. Nearly all of our Barium candidates are disk stars, 178 exhibit CH and G band enhancement, no plausible technetium enhancement is discovered, and it is shown sodium enhancement is not ubiquitous in s-process stars as previously claimed. \todo{AGB yields comparison. Something about binary eccentricities.}
\section*{Acknowledgements}

Fired Status: yes

@@

@@

@@

@@

\bibitem[\protect\citeauthoryear{Casey}{2014}]{casey2014}
Casey, A.~R. 2014, 
PhDT, 394
\todo{format Andy's ref properly}
%@PHDTHESIS{2014PhDT.......394C,
%   author = {{Casey}, A.~R.},
%    title = "{A Tale of Tidal Tales in the Milky Way}",
% keywords = {Astronomy, Astrophysics, Stellar streams},
%   school = {Australian National University},
%     year = 2014,
%    month = May,
%      doi = {10.5281/zenodo.49493},
%   adsurl = {http://adsabs.harvard.edu/abs/2014PhDT.......394C},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%}

>>>>

\begin{abstract}
Stars with peculiar enhancements of carbon and heavy elements ($Z > 30$) are historically referred to as `Barium stars'. While this chemical abundance pattern can be explained by thermally pulsing asymptotic giant branch (AGB) stars, many so-called Barium stars have not evolved through the red giant phase and therefore cannot be AGB stars. These less-evolved stars are generally thought to have accreted mass from a more-evolved AGB companion. For this reason the frequency and properties of less-evolved Barium stars are uniquely informative on the binary fraction, AGB yields, mass transfer, and stellar evolution. Here we present the discovery of 967 s-process candidates from 454,180 giant stars observed by LAMOST. Our sample includes; 257 Barium stars, 49 stars exhibiting only significant barium enhancement, and 661 exhibiting only strontium enhancement. This sample constitutes the largest number of s-process enhanced stars ever discovered, and represents the most comprehensive search for Barium stars in the Milky Way. All of our Barium stars are consistent with membership in the Milky Way disk, and we find an \todo{steady/increasing/decreasing} fraction of Barium stars with stellar metallicity, as \todo{expected} from AGB yields and stellar evolution theory. We find up to 51 Barium stars exhibit some evidence of technetium from LAMOST spectra, and most (46/51) have $\log{g} > 2$, which suggests they cannot be AGB stars and may indicate a recent ($<1\,\textrm{Myr}$) mass accretion event. Most (178/257) Barium stars in our sample show strong carbon enhancement, consistent with the literature. However, despite previous claims indicating that Barium stars are preferentially enhanced in sodium due to the NeNa nucleosynthesis cycle occurring in massive AGB stars, only 5 ($<2$\,\%) of our Barium stars show any evidence of enhancement in sodium. An AGB yield comparison demonstrates that the majority of our sample agree with the Salpeter IMF, and highlights a small percentage of intrinsic Barium stars present in our sample. When binary properties become available from end-of-mission Gaia astrometry, this sample of s-process stars will be critical in explaining mass transfer in binary systems and refining AGB yields across a range of masses and metallicities.
\end{abstract}

@@

\subsection{Extrinsic and intrinsic Barium stars}
Barium stars exist as either intrinsic or extrinsic objects. Intrinsic Barium stars must be massive enough to reach the TP-AGB phase, and extrinsic Barium stars must be in a binary with a previously polluting TP-AGB star companion. Figure \ref{fig:figure2} shows that the majority of our Barium stars have $\log{g} > 2$, highlighting that most stars in our sample are not massive enough to have already reached the TP-AGB phase and therefore, must be extrinsic Barium stars. The classification for Barium stars with $\log{g} < 2$ is uncertain: these stars could be intrinsic or extrinsic. From here on when we refer to extrinsic Barium stars, we specifically refer to those with $\log{g} > 2$. It is noted, in the previous literature \citet{van2017}, only 22\% of their stellar sample are post-main-sequence barium stars (extrinsic barium stars). In comparison, we find our Barium candidates constitute greater than 70\% extrinsic, this suggests the construction of our data-driven model may have missed some AGB Barium candidates.

\subsection{Sodium enhancment}
According to \citet{el1995}, sodium overabundances observed in the atmospheres of supergiants is synthesised by the NeNa reaction chain in the convective core of main-sequence stars. Sodium is transported to the surface of these supergiants via mixing of CNO cycle products during the first dredge-up, although, it is important to note sodium enhancement is not constrained to supergiant stars. Stars with a  minimum mass of $1.5\,M_\odot$ can also exhibit enhanced sodium abundances through the NeNa cycle \citep{denissenkov1987}. \citet{antipova2004} reported sodium enhancements in 3 of their 16 Barium star candidates, and they relate this to the dredge-up of nuclear-burning material produced by convection during the red-giant phase. Moreover, they suggest [Na/Fe] ratios exhibit systematically higher values for giants with lower $\log{g}$ values, but do not exhibit variations in metallicity. Similarly, \citet{decastro2016} highlights the possible weak anti-correlation between [Na/Fe] ratio and $\log{g}$, and highlights that this trend in previous studies \citep[e.g.][]{boyarchuk2002,mishenina2006,luck2007,takeda2008}.

We find only $<2$\,\% of our Barium stars show sodium enhancement, contrary to previous studies \citep[e.g.][]{decastro2016}. We also note that the 5 sodium-enhanced Barium stars all have $\log{g} \approx 3$, not at lower $\log{g}$ values as might be expected if low $\log{g}$ values systematically contributed to higher [Na/Fe] abundance ratios. These results suggest that sodium-rich material from a companion TP-AGB star has polluted these 5 Barium candidates, but importantly, sodium enhancement is by no means ubiquitous in Barium stars. It's important to note, sodium enhancement can occur due to the presence of interstellar dust, and using the IRSA's all-sky dust map \citep{schlafly2011}, E(B-V) values were calculated for our 5 candidates. The results show 2 sodium enhanced stars have strong ($\approx$0.35) E(B-V) values, higher then galactic stellar levels, suggesting these candidates are false positives. Whereas, 3 exhibit low ($\approx$0.045) E(B-V) values, validating our sodium enhancement results for these candidates. The comparisons made in \citep{decastro2016} made use of multiple literature comparisons, and therefore it is possible that there are systematic differences (biases) between Milky Way studies of normal FGK-type stars and those focussed on Barium stars, leading to unresolved differences in the reported [Na/Fe] abundance ratio. These same systematic effects could contribute to the observed correlation between [Na/Fe] and $\log{g}$, which we do not observe in our sample.

\subsection{Technetium enhancement}

Technetium has a half life of approximately 211,000 years and as a result, Tc enhacnement is constrained to observations intrinsic Barium stars \citep{jorissen1993} or recently polluted extrinsic Barium stars. Our analyse discovered 51 single Tc line matches within our 257 Barium star candidate sample. The majority (46/51) have $\log{g} > 2$, suggesting they cannot be AGB stars, and may indicate a recent ($<1\,\textrm{Myr}$) mass accretion event. However, for Tc to be present in  a star the enhancement lines must be considerably strong, and given that all 51 matches were at single lines it is concluded that these results are more likely data artifacts. In agreement with previous findings \citep[e.g.][]{little1987,smith1984,smith1983}, the results indicate that none of the evolved Barium stars in our sample exhibited the presence of Tc enhancement, and it can be concluded our Barium star sample is extrinsic. It is noted, Tc enhancement is a very difficult line to analyse and that this conclusion is not concrete.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{HRdummy.png}
    \caption{Effective temperature $T_{eff}$ and surface gravity $\log{g}$ for 257 candidate Barium stars from LAMOST. Stars with $\log{g} > 2$ (as marked) have not evolved through the thermally pulsing asymptotic giant branch phase, and therefore their abundance signature must be the result of an extrinsic pollution event. The classification of stars with $\log{g} < 2$ would require additional data}
    \label{fig:figure2}
\end{figure}

\subsection{AGB yields comparison}
AGB stars, and chemically unique stars that show the chemical signature
of mass transfer pollution from binary companion AGB stars, highlight the variation of s-process enhancement. \citet{karakas_lugaro2016} demonstrates through stellar yield models, that Barium stars will have a likely minimum yeild at [Fe/H]$\approx$ -1.2. Figure \ref{fig:figure3} confirms this as the majority of the Barium stars are contained above -1.2. The final surface s-process index [hs/ls], shown in \citet{cristallo2015}, highlights the relationship between [hs/ls] yields and the mass of AGB stars. The Salpeter IMF predicts the majority of stars in the Milky Way will be at lower mass, and as predicted the [hs/ls] values of our Barium stars approximately fit the lower mass yields shown in \citet{cristallo2015}. Barium stars not fitting this relationship represent the possibly more massive AGB stars (intrinsic Barium stars) within our sample.

\begin{figure}
	\includegraphics[width=\columnwidth]{basrtest4.png}
    \caption{Final surface s-process indexes: [hs/ls]}
    \label{fig:figure3}
\end{figure}

@@

Our research discovered 967 s-process candidates including; 257 Barium stars, 49 stars exhibiting exclusively barium enhancement, and 661 exhibiting exclusively strontium enhancement from 454,180 giant stars observed by LAMOST. This is double the currently known stellar sample of Barium stars, and the largest sample of s-process enhanced stars known to date. Nearly all of our Barium candidates are disk stars, 178 exhibit CH and G band enhancement, no plausible technetium enhancement is discovered, and it is shown sodium enhancement is not ubiquitous in s-process stars as previously claimed. AGB yields comparisons indicate the majority of our sample fit low steallar mass trends in agreement with Salpeter IMF theory. Additonally, a number of outliers highlight a small percentage of intrinsic Barium stars in our sample.
todo{ Something about binary eccentricities.}
\section*{Acknowledgements}

Fired Status: maybe

@@

\bibitem[\protect\citeauthoryear{Casey}{2014}]{casey2014}
Casey, A.~R. 2014, 
PhD Thesis, Australian National University
\bibitem[\protect\citeauthoryear{Cristallo et al.}{2015}]{cristallo2015}
Cristallo, S., Straniero, O., Piersanti, L.,\& Gobrecht, D. 2015, 
ApJ, 219, 40

@@

\bibitem[\protect\citeauthoryear{Karakas \& Lugaro}{2016}]{karakas_lugaro2016}
Karakas, A.~I.,\& Lugaro M. 2016, 
ApJS, 825, 26

@@

\bibitem[\protect\citeauthoryear{Schlafly \& Finkbeiner}{2011}]{schlafly2011}
Schlafly, E.~F.,\& Finkbeiner, D.~P. 2011, 
Apj, 737, 103

@@

\bibitem[\protect\citeauthoryear{Van der Swaelmen et al.}{2017}]{van2017}
Van der Swaelmen, M., Boffin, H.~M.~J., Jorissen, A.,\& Van Eck, S. 2017, 
A$\&$A, 597, A68
